Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-13 surfaces-write-up/surface-write-up.tex
lemma 5.3 edits

Let $E$ be the set of $\epsilon_i$ and $\epsilon_{i,j}$ given 
above, satisfying the respective conditions.
$E$ satisfies the conditions of proposition \todo{cite earlier 
proposition}, so by applying that proposition $R_D$ is generated in 
degree at most
\[
	\sum_{i\in T_=(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{i\in T_+(D) \atop
	j \in T_-(D)} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\end{proof}

@@

\begin{proof}
We first claim that $\psi$ has a relation of the form
\begin{equation}\label{eqn:relations-psi}
	(y_j - \beta_j)\prod_{i=1}^n {y_i}^{c_{i}}
\end{equation}

@@

Note that we can view the relations from Equation
\ref{eqn:relations-psi} as an element of $S$ given
by $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$.  Let $S'\subseteq (\mathbb{R}_{\ge 0})^{n+1}$ be the
set of $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$ satisfying $c_i \ge -\alpha_i d$,
$\bida_j + \sum_{i = 1}^m \bida_i c_i = 0$, and $\bidb_j + \sum_{i=1}^n \bidb_i c_i
= 0$.  A careful computation deduces that $\gamma (d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots, c_n)$ is of the form 
\[
	E + \sum_{k\in T_=} r_k e_k + \sum_{h\in T_+ \atop k\in T_-} r_{h,k} e_{h,k},
\] 
where $E$ is $\mathbb{R}_{\ge 0}$
combination of elements in 
\[
	\{e_i \mbox{ if } i\in S\} \cup \{e_{i,j}: j\in S_- \mbox{ if } i \in S_+\} \cup \{e_{j,i}: j\in S_+ \mbox{ if } i \in S_-\}
\]
given by lattice point $(d_E, w_1, \ldots, w_n)$ with $w_i = 1$.  These conditions imply that 
\[
	d_E \le \left\{ \begin{split}  \ell_i \gcd(\bida_i, \bidb_i) \mbox{ if } i \in S_= 
			\\ \max_{j\in S_-}(\ell_{i,j} (\bida_i \bidb_
j - \bida_j \bidb_i)) \mbox{ if } i\in S_+ \\
			\\ \max_{j\in S_+}(\ell_{j, i} (\bida_j \bidb_
i - \bida_i \bidb_j)) \mbox{ if } i \in S_- \end{split}  \right.
\]

Let $M$ be the set of monomials of $\beta_j$.  Then for $m\in M$, the corresponding lattice point is of the form
\[
	E_m + \sum_{k\in T_=} r_k e_k + \sum_{h\in T_+ \atop k\in T_-} r_{h,k} e_{h,k}
\]
where $E_m$ is a $\mathbb{R}_{\ge 0}$ combination of elements in
\[
	\{e_j : j\in T_= \cap \{1, 2, 3, 4\}\} \cup \{e_{j, k}: j\in T_= \cap \{1, 2, 3, 4\}, k\in T_-\} \cup \{e_{k,j}: k\in T_+, j\in T_- \cap \{1, 2, 3, 4\}\}.
\]

>>>>

Let $E$ be the set of $\epsilon_i$ and $\epsilon_{i,j}$ given 
above, satisfying the respective conditions.
$E$ satisfies the conditions of proposition \todo{cite earlier 
proposition}, so by applying that proposition $R_D$ is generated in 
degree at most
\[
	\sigma = \sum_{i\in T_=(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{i\in T_+(D) \atop
	j \in T_-(D)} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\end{proof}

@@

\begin{proof}
We first claim that $\psi$ has a relation of the form
\begin{equation}\label{eqn:relations-psi}
	(z_i - \beta_j)\prod_{j=1}^n {z_j}^{c_{i}}
\end{equation}

@@

Note that we can view the relations from Equation
\ref{eqn:relations-psi} as an element of $S$ given
by $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$.  Let $S'\subseteq (\mathbb{R}_{\ge 0})^{n+1}$ be the
set of $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$ satisfying $c_i \ge -\alpha_i d$,
$\bida_j + \sum_{i = 1}^m \bida_i c_i = 0$, and $\bidb_j + \sum_{i=1}^n \bidb_i c_i
= 0$.  A careful computation deduces that $\gamma (d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots, c_n)$ is of the form 
\[
	E + \sum_{k\in T_=} r_k e_k + \sum_{h\in T_+ \atop k\in T_-} r_{h,k} e_{h,k},
\] 
where $E$ is $\mathbb{R}_{\ge 0}$
combination of elements in 
\[
	\{e_i \mbox{ if } i\in S\} \cup \{e_{i,j}: j\in S_- \mbox{ if } i \in S_+\} \cup \{e_{j,i}: j\in S_+ \mbox{ if } i \in S_-\}.
\]

Let $M$ be the set of monomials of $\beta_j$.  Then for $\mu\in M$, the corresponding lattice point is of the form
\[
	E_\mu + \sum_{k\in T_=} r_k e_k + \sum_{h\in T_+ \atop k\in T_-} r_{h,k} e_{h,k}
\]
where $E_\mu$ is a $\mathbb{R}_{\ge 0}$-combination of elements in
\begin{align*}
			& \{e_j : j\in T_= \cap \{1, 2, 3, 4\}\} \\
	\cup \; 	& \{e_{j, k}: j\in T_= \cap \{1, 2, 3, 4\}, k\in T_-\}\\
	\cup \; 	& \{e_{k,j}: k\in T_+, j\in T_- \cap \{1, 2, 3, 4\}\}.
\end{align*}

Let $F\in R_D$ correspond to $E + \sum_{k\in T_=} (r_k) e_k + \sum_{h\in T_+ \atop k\in T_-} (r_{h,k}) e_{h,k}$ where $(r)$ signifies the fractional part of $r$.  Similarly, let $F_\mu \in R_D$ correspond to $E_\mu + \sum_{h\in T_+ \atop k\in T_-} (r_{h,k}) e_{h,k}$.  Then $F - \sum_{\mu\in M} F_\mu = 0$ gives us a relation in degree at most
\[
	\sigma + \deg(F),
\]
which generates $(z_i - \beta_i) \prod_{i=1}^n y_i^{c_i}$
We can bound $\deg(F)$ by
\[
	\deg(F) \le \left\{ \begin{split}  \ell_i \gcd(\bida_i, \bidb_i) \mbox{ if } i \in S_= 
			\\ \max_{j\in S_-}(\ell_{i,j} (\bida_i \bidb_
j - \bida_j \bidb_i)) \mbox{ if } i\in S_+ \\
			\\ \max_{j\in S_+}(\ell_{j, i} (\bida_j \bidb_
i - \bida_i \bidb_j)) \mbox{ if } i \in S_- \end{split}  \right.
\]
Therefore, $\ker(\psi)$ is generated in degree at most
\[
	\tau = \sigma
	+ \max( \max_{i\in S_=}(\ell_i \gcd(a_i, b_i)), \max_{i\in S_+ \atop j\in S_-}(\ell_{i,j} (\bida_i \bidb_j - \bida_j \bidb_i))).
\]
